\documentclass[10pt]{article}
\usepackage{../pplmanual}
\input{../pplmanual}

\title{BigSimulator (BigNetSim) for Extremely Large Parallel Machines}
\version{1.01}
	\credits{The Charm++ BigSim Emulator was developed by Arun Singla,
Neelam Saboo and Joshua Unger under the guidance of Prof. L. V. Kale. The new
Converse BigSim Emulator was completely rewritten by Gengbin Zheng. The
Converse BigSim Emulator is the only version under maintenance now. Charm++ and
Adaptive MPI (AMPI) were ported onto the BigSim Emulator by Gengbin Zheng. The
parallel performance simulator was developed by Gengbin Zheng and Gunavardhan
Kakulapati. A postmortem network simulator was developed by Terry Wilmarth,
Eric Bohm and Gengbin Zheng}

\begin{document}
\maketitle

\section{BigSim Network Simulator}
\label{bignetsim}

The BigSim Network Simulator is also known as Bigsimulator and lives
in the SVN repository https://charm.cs.uiuc.edu/svn/repos/BigNetSim.
The Network simulator is actually more of an Inter-connection network
simulator and hence more important in the context of large parallel
machines with interconnects.
The BigSim simulator  along with the network simulator is together
also known as BigNetSim.

Both the simulators run on top of the POSE framework, which is a Parallel
Discrete Event Simulation framework built on top of \charmpp{}.


\subsection{What does this software do?}
BigNetSim is an effort to simulate large current and future computer
systems to study the behavior of applications developed for those systems.
BigNetSim could be used to study
\begin{itemize}
\item  new types of interconnection topologies and routing algorithms
along with different types of switching architecture.
\item application performance on different machines. This uses the API
provided in Section \ref{bgapi} to run the application on some number
of processors on some machine and generate (dump) all events (entry
method executions or message send/recv).  BigNetSim is used to
model the machine that needs to be studied for this application and
these logs are then fed into this simulation, and it predicts the
performance of this application.
\end{itemize}

So, the two important uses are studying {\it interconnection networks} and
{\it performance prediction for applications}.

\input{install}

\input{usage}

\input{interconnects}

\input{bignetsim}

\input{index}

\end{document}

