\subsection{Entry Methods}

\label{entry}

In \charmpp, \index{chare}chares, \index{group}groups and \index{nodegroup}
nodegroups communicate using remote method invocation.  These ``remote entry'' methods may either take marshalled parameters, described in the next section; or special objects called messages.  Messages are lower level, more efficient, more flexible, and more difficult to use than parameter marshalling.

An entry method is always a part of a chare--
there are no global entry methods in \charmpp{}.
Entry methods are declared in the the interface file as:

\begin{alltt}
entry void \uw{Entry1}(\uw{parameters});
\end{alltt}

\uw{Parameters} is either a list of marshalled parameters,
(e.g., ``int i, double x''), or a message description (e.g.,
``MyMessage *msg'').  See section~\ref{marshalling} and
section~\ref{messages} for details on these types of
parameters.

Entry methods typically do not return data-- in \CC, they have
return type ``void''.  An entry method with the same name
as its enclosing class is a constructor.  Constructors in \CC
have no return type.  Finally, sync methods, described below,
may return a message.

\subsection{Entry Method Invocation}

\label{marshalling}

\experimental{}	
In \charmpp, \index{chare}chares, \index{group}groups and \index{nodegroup}
nodegroups communicate by invoking each others methods. 
The methods may either take several parameters, described here; 
or take a special message object as described in the next section.
Since parameters get marshalled into a message before being
sent across the network, in this manual we use ``message''
to mean either a literal message object or a set of marshalled
parameters.

For example, a chare could have this entry method declaration in 
the interface ({\tt .ci}) file:
\begin{alltt}
  entry void foo(int i,int k);
\end{alltt}
Then invoking foo(2,3) on the chare proxy will eventually
invoke foo(2,3) on the remote chare.

Since \charmpp\ runs on distributed memory machines, we cannot
pass an array via a pointer in the usual \CC\ way.  Instead,
we must specify the length of the array in the interface file, as:
\begin{alltt}
  entry void bar(int n,double arr[n]);
\end{alltt}
Since \CC\ does not recognize this syntax, the array data
must be passed to the chare proxy as a simple pointer.
The array data will be copied and sent to the
destination processor, where the chare will receive the copy
via a simple pointer again.  The remote copy of the data
will be kept until the remote method returns, when
it will be freed.  
This means any modifications made locally after the call will not be 
seen by the remote chare; and the remote chare's modifications
will be lost after the remote method returns-- \charmpp\ always 
uses call-by-value, even for arrays and structures.  

This also means the data must be copied on the sending 
side, and to be kept must be copied again 
at the receive side.  Especially for large arrays, this 
is less efficient than messages, as described in the next section.

Array parameters and other parameters can be combined in arbitrary ways, as:
\begin{alltt}
  entry void doLine(float data[n],int n);
  entry void doPlane(float data[n*n],int n);
  entry void doSpace(int n,int m,int o,float data[n*m*o]);
  entry void doGeneral(int nd,int dims[nd],float data[product(dims,nd)]);
\end{alltt}
The array length expression between the square brackets can be 
any valid C++ expression, including a fixed constant, and may depend 
in any manner on any of the passed
parameters or even on global functions or global data.  The array length 
expression is evaluated exactly once per invocation, on the sending side only.
Thus executing the \kw{doGeneral} method above will invoke the 
(user-defined) \kw{product} function exactly once on the sending
processor.

\subsubsection{Marshalling User-Defined Structures and Classes}

The marshalling system uses the pup framework to copy data,
meaning every user class that is marshalled needs either a
pup routine, a ``PUPbytes'' declaration, or a working operator|.
See the PUP description in Section~\ref{sec:pup} for more details 
on these routines.

\begin{alltt}
//Declarations:
class point3d \{
public:
    double x,y,z;    
    void pup(PUP::er &p) \{
      p|x; p|y; p|z;
    \}
\};

typedef struct \{
    int refCount;
    char data[17];
\} refChars;
PUPbytes(date);

class date \{
public:
    char month,day;
    int year;
    //...non-virtual manipulation routines...
\};
inline void operator|(PUP::er &p,date &d) \{
    p|d.month; p|d.day;
    p|d.year;
\}

//In the .ci file:
    entry void pointRefOnDate(point3d &p,refChars r[d.year],date &d);
\end{alltt}

Any user-defined types in the argument list must be declared 
before including the ``.decl.h'' file.
As usual in \CC, it is often dramatically more efficient to pass
a large structure by reference (as shown) than by value.

For efficiency, arrays (like \uw{refChars} above) are always copied
as blocks of bytes and passed via pointers.  This means classes
that need their pup routines to be called, such as those with dynamically
allocated data or virtual methods cannot be passed as arrays--use CkVec
or STL vectors to pass lists of complicated user-defined classes. 
For historical reasons, pointer-accessible structures 
cannot appear alone in the parameter list (because they are confused 
with messages).

The order of marshalling operations on the send side is:
\begin{itemize}
\item Call ``p\verb.|.a'' on each marshalled parameter with a sizing PUP::er.
\item Compute the lengths of each array.
\item Call ``p\verb.|.a'' on each marshalled parameter with a packing PUP::er.
\item \kw{memcpy} each arrays' data.
\end{itemize}

The order of marshalling operations on the receive side is:
\begin{itemize}
\item Create an instance of each marshalled parameter using its default constructor.
\item Call ``p\verb.|.a'' on each marshalled parameter using an unpacking PUP::er.
\item Compute pointers into the message for each array.
\end{itemize}

Finally, very large structures are most efficiently passed via messages,
because messages are an efficient, low-level construct that minimizes copying
and overhead; but very complicated structures are easiest to pass via 
marshalling, because marshalling uses the high-level pup framework.
