\section{Utility Functions}

\label{basic utility fns}

The following calls provide basic rank information and utilities
useful when running a Charm++ program.

\function{void CkAssert(int expression)} \desc{Aborts the program 
if expression is 0.}

\function{void CkAbort(const char *message)} \index{CkAbort}
\desc{Causes the program to abort, printing the given error message.
  This function never returns.}

\function{void CkExit()} \index{CkExit} \desc{This call informs the
  Charm RTS that computation on all processors should terminate.  This
  routine never returns, so any code after the call to CkExit() inside
  the function that calls it will not execute. Other processors will
  continue executing until they receive notification to stop, so it is
  a good idea to ensure through synchronization that all useful work
  has finished before calling CkExit(). Application code should be
  structured so that it makes only a single call to CkExit() across
  the entire system, such as by selecting a designated object to make
  the call. Multiple calls can lead to undefined behavior. }

\function{double CkWallTimer()} \index{CkWallTimer} \index{timers}
\desc{Returns the elapsed time in seconds since the start of the program.}

\subsection{Information about Logical Machine Entities}
As described in section~\ref{machineModel}, \charmpp{} recoginizes two
logical machine entities: ``node'' and PE (processing element). 
The following functions provide basic information about such logical machine 
that a \charmpp{} program runs on. PE and ``node'' are numbered starting
from zero.

\function{int CkNumPes()} \index{CkNumPes} \desc{returns the total
  number of PEs across all nodes.}

\function{int CkMyPe()} \index{CkMyPe} \desc{returns the index of the
  PE on which the call was made.}

\function{int CkNumNodes()} \index{CkMyNodes}
\desc{returns the total number of logical \charmpp{} nodes.}

\function{int CkMyNode()} \index{CkMyNode}
\desc{returns the index of the ``node'' on which the call was made.}

\function{int CkMyRank()} \index{CkMyRank}
\desc{returns the rank number of the PE on a ``node'' on which the call 
was made. PEs within a ``node'' are also ranked starting from zero.}

\function{int CkNodeFirst(int nd)} \index{CkNodeFirst}
\desc{returns the index of the first PE on the logical node $nd$.}

\function{int CkNodeSize(int nd)} \index{CkNodeSize}
\desc{returns the number of PEs on the logical node $nd$ on which the call was made.}

\function{int CkNodeOf(int pe)} \index{CkNodeOf}
\desc{returns the ``node'' number that PE $pe$ belongs to.}

\function{int CkRankOf(int pe)} \index{CkRankOf}
\desc{returns the rank of the given PE within its node.}

\subsection{Terminal I/O}

\index{input/output}
\charmpp\ provides both C and \CC\ style methods of doing terminal I/O.  

In place of C-style printf and scanf, \charmpp\ provides
\kw{CkPrintf} and \kw{CkScanf}.  These functions have
interfaces that are identical to their C counterparts, but there are some
differences in their behavior that should be mentioned.

\charmpp\ also supports all forms of printf,
cout, etc. in addition to the special forms shown below.  The special
forms below are still useful, however, since they obey well-defined
(but still lax) ordering requirements.

\function{int CkPrintf(format [, arg]*)} \index{CkPrintf}
\index{input/output} \desc{This call is used for atomic terminal
  output. Its usage is similar to \texttt{printf} in C.  However,
  \kw{CkPrintf} has some special properties that make it more suited
  for parallel programming.  \kw{CkPrintf} routes all terminal output
  to a single end point which prints the output. This guarantees that
  the output for a single call to CkPrintf will be printed completely
  without being interleaved with other calls to CkPrintf. Note that
  CkPrintf is implemented using an asynchronous send, meaning that the
  call to \kw{CkPrintf} returns immediately after the message has been
  sent, and most likely before the message has actually been received,
  processed, and displayed. As such, there is no guarantee of order in
  which the output for concurrent calls to CkPrintf is
  printed. Imposing such an order requires proper synchronization
  between the calls to CkPrintf in the parallel application.
}

\function{void CkError(format [, arg]*))} \index{CkError} \index{input/output} 
\desc{Like \kw{CkPrintf}, but used to print error messages on \texttt{stderr}.}

\function{int CkScanf(format [, arg]*)} \index{CkScanf} \index{input/output}
\desc{This call is used for atomic terminal input. Its usage is similar to
{\tt scanf} in C.  A call to \kw{CkScanf}, unlike \kw{CkPrintf},
blocks all execution on the processor it is called from, and returns
only after all input has been retrieved.
}

For \CC\ style stream-based I/O, \charmpp\ offers 
\kw{ckout} and \kw{ckerr} in place of cout and cerr.  The
\CC\ streams and their \charmpp\ equivalents are related in the same
manner as printf and scanf are to \kw{CkPrintf} and \kw{CkScanf}.  The
\charmpp\ streams are all used through the same interface as the \CC\ 
streams, and all behave in a slightly different way, just like C-style
I/O.
